---
title: "Final Project"
author: "Matthew Roland"
date: "2023-12-07"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(ggplot2)
library(GGally)
library(Matrix)
library(pracma)
library(matrixcalc)
library(MASS)
library(car)
```


**Modeling.  Build some type of multiple regression  model and submit your model to the competition board.  Provide your complete model summary and results with analysis.  Report your Kaggle.com user name and score.**

```{r}
crab_train <- read.csv("https://raw.githubusercontent.com/Mattr5541/CUNY-School-of-Professional-Studies-Project-Archive/main/Regression%20with%20a%20Crab%20Kaggle%20Competition/train.csv")

crab_sex <- crab_train %>% count(Sex) %>% reframe(Sex = Sex, count = n, prop = n / sum(n))

kable(crab_sex)

crab_sex %>% ggplot(aes(Sex, prop)) + geom_bar(stat = "identity")
```

As we can see, there is a larger proportion of males compared to indeterminate or female crabs in this dataset; however, the ratios are rather similar, overall.

```{r}
crab_mean <- crab_train %>% summarize(length = mean(Length), 
                                      diameter = mean(Diameter),
                                      height = mean(Height),
                                      shucked_weight = mean(Shucked.Weight),
                                      viscera_weight = mean(Viscera.Weight),
                                      shell_weight = mean(Shell.Weight),
                                      age = mean(Age),
                                      weight = mean(Weight)) %>% mutate(descriptive = "mean")

crab_median <- crab_train %>% summarize(length = median(Length), 
                                      diameter = median(Diameter),
                                      height = median(Height),
                                      shucked_weight = median(Shucked.Weight),
                                      viscera_weight = median(Viscera.Weight),
                                      shell_weight = median(Shell.Weight),
                                      age = median(Age),
                                      weight = median(Age)) %>% mutate(descriptive = "median")

crab_sd <- crab_train %>% summarize(length = sd(Length), 
                                      diameter = sd(Diameter),
                                      height = sd(Height),
                                      shucked_weight = sd(Shucked.Weight),
                                      viscera_weight = sd(Viscera.Weight),
                                      shell_weight = sd(Shell.Weight),
                                      age = sd(Age),
                                      weight = sd(Weight)) %>% mutate(descriptive = "sd")

crab_table <- bind_rows(crab_mean, crab_median, crab_sd)

kable(crab_table)

crab_table_long <- crab_table %>% pivot_longer(!descriptive, names_to = "features")

crab_table_long %>% filter(descriptive == "mean") %>% ggplot(aes(features, value)) + geom_bar(stat = "identity")

crab_train_long <- crab_train %>% pivot_longer(!c(id, Sex, Age), names_to = "features")

crab_train_long %>% filter(features == c("Diameter", "Height", "Length")) %>% ggplot(aes(value)) + geom_histogram() + facet_wrap(vars(features))

crab_train_long %>% filter(features == c("Shucked.Weight", "Viscera.Weight", "Shell.Weight", "Weight")) %>% ggplot(aes(value)) + geom_histogram() + facet_wrap(vars(features))

crab_train %>% ggplot(aes(Age)) + geom_histogram()
```

As we can see, most of the variables in this dataset have some form of positive or negative skew

```{r}
crab_test <- read.csv("https://raw.githubusercontent.com/Mattr5541/CUNY-School-of-Professional-Studies-Project-Archive/main/Regression%20with%20a%20Crab%20Kaggle%20Competition/test.csv")

crab_reg <- lm(Age ~ Shell.Weight * Shucked.Weight * Weight, crab_train)

summary(crab_reg)

plot(crab_reg)
```

As we can see, there are some serious violations in the assumptions for our regression analysis, so I will remove outliers for each outcome and predictor by using the interquartile range


```{r}
outliers_Age <- boxplot(crab_train$Age, plot = F)$out
outliers_Weight <- boxplot(crab_train$Weight, plot = F)$out
outliers_Shuck <- boxplot(crab_train$Shucked.Weight, plot = F)$out
outliers_Shell <- boxplot(crab_train$Shell.Weight, plot = F)$out




crab_train_mod <- crab_train
crab_train_mod <- crab_train_mod[-which(crab_train_mod$Age %in% outliers_Age),]


crab_train_mod %>% ggplot(aes(Age)) + geom_histogram()
crab_train_mod %>% ggplot(aes(Shell.Weight)) + geom_histogram()
crab_train_mod %>% ggplot(aes(Shucked.Weight)) + geom_histogram()
crab_train_mod %>% ggplot(aes(Weight)) + geom_histogram()

crab_reg2 <- lm(Age ~ Shell.Weight * Shucked.Weight * Weight, crab_train_mod) 
summary(crab_reg2)

plot(crab_reg2)

crab_pred <- predict(crab_reg2, crab_test )

crab_pred <- as.data.frame(crab_pred)

crab_pred <- cbind(crab_test, crab_pred) %>% rename(Age = "crab_pred")

crab_pred <- crab_pred %>%  dplyr::select(c(id, Age))

write.csv(crab_pred, "C:\\Users\\Matt\\Documents\\605\\crab_pred.csv", row.names = F)

```

Kaggle username: Matt Roland
Kaggle Score: 1.48459
